---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 
library(gghighlight)
library(geomtextpath)
```

### Exercise 1 & 2

```{r employ, width = 10}
staff <- read_csv("data/instructional-staff.csv")

staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))

staff_long %>%
  ggplot(aes(x = year,
             y = value,
             group = faculty_type,
             color = faculty_type))+
  geom_line()+
   gghighlight(faculty_type == "Part-Time Faculty", use_direct_label = FALSE, unhighlighted_params = aes(color = NULL, alpha = 0.2), keep_scales = TRUE)+
  labs(
    title = "Trends in Instructional Staff Employment Status, 1975-2011",
    subtitle = "All Institutions, National Totals",
    y = "Percent of Total Instructional Staff"
  )+
  scale_color_discrete(name  ="faculty type")
# I highlighted the part-time compared to others so that it is more salient in the graph. 

```


### Exercise 3

I want to find a better way to filter the country variable so that it looks cleaner the plot. Also, I want to overlap two variables in the same plot for the easy of comparison. 

```{r Fisheries, width = 5}
fisheries <- read_csv("data/fisheries.csv")

fisheries %>%
  filter (total > 1000000) %>%
  ggplot()+
  geom_col(aes(y=reorder(country,total), x= total,fill = "total"))+
  geom_col(aes(y=reorder(country,total), x= aquaculture,fill = "aquaculture"))

fisheries %>%
  filter (total > 1000000) %>%
  ggplot()+
  geom_col(aes(y=reorder(country,total), x= total,fill = "total"))+
  geom_col(aes(y=reorder(country,total), x= capture,fill = "capture"))

fisheries %>%
  filter (total > 1000000) %>%
  ggplot()+
  geom_col(aes(y=reorder(country,total), x= capture,fill = "capture"))+
  geom_col(aes(y=reorder(country,total), x= aquaculture,fill = "aquaculture", alpha = 0.5))




```

### Exercise 1 (Wickham)

This is observational data because there is no experimental manipulation in this study. 

```{r data_input, echo=FALSE}
whickham <- force(data(Whickham))
```


### Exercise 2 (Wickham)

There are 1314 pieces of observation and they each represent a single individual. 
```{r whickham_dim}
dim(Whickham)
head(Whickham)
```

### Exercise 3 (Wickham)

There are 3 variables. Outcome and smoker are discret variables while age is a continuous variable.

```{r whickham_vars}
ncol(Whickham)

Whickham %>%
  ggplot(aes(x=outcome, fill = smoker))+
  geom_bar()

Whickham %>%
  ggplot(aes(x=age))+
  geom_density()
```

### Exercise 4 (Wickham)

I would expect the relationship between smoking status and health outcome to be positive.

```{r cor}
chisq.test(Whickham$outcome,Whickham$smoker)
```


### Exercise 5 (Wickham)

```{r age_cat}

Whickham$age_cat <- ifelse(Whickham$age <= 44, "18-44", 
                           ifelse((Whickham$age > 44& Whickham$age <= 64),"45-64","65+"))

```

### Exercise 6 (Wickham)

```{r age_dif}
Whickham %>%
  count(smoker, age_cat, outcome)%>%
  ggplot(aes(x=outcome, y = n, fill = smoker))+
  geom_bar(stat = "identity")+
  facet_grid(cols = vars(age_cat))
```

